Bayesian Exploration in Deep Reinforcement Learning

### 2022-01-28 papers/main.tex
TEMPORARY DEPLOY COMMIT

%% Some suggested packages, as needed:
\usepackage{natbib} % has a nice set of citation styles and commands
    \bibliographystyle{plainnat}
    \renewcommand{\bibsection}{\subsubsection*{References}}
\usepackage{mathtools} % amsmath with fixes and additions
% \usepackage{siunitx} % for proper typesetting of numbers and units
\usepackage{booktabs} % commands to create good-looking tables
\usepackage{tikz} % nice language for creating drawings and diagrams
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{bm}
\usepackage{subcaption}

@@

In this paper we will introduce an extension to NoisyNet that puts it into a Bayesian context. A limitation of NoisyNet is that the initial uncertainty in the Q-value function is crucial to exploration. If the uncertainty is too high, algorithm will struggle to learn anything, while if the uncertainty is too low, there is nothing incentivizing exploration, and the algorithm will likely be stuck in a poor local minimum.

@@

\subsection{Noisy Networks}
In NoisyNet \citep{fortunato_noisy_2019}, the Q-function space can be seen
as a stochastic process on the probability space \((\Omega, \mathcal{F}, P)\).
The stochastic process can simply be written as
\(\{Q(\bm{s}, \bm{a}) : (\bm{s}, \bm{a}) \in \state \times \action \}\). For
any \(\omega \in \Omega\), \(Q(\cdot, \cdot, \omega)\) is a sample function
mapping \(\state \times \action \rightarrow \mathbb{R}\). To simplify notation
we will denote sample functions as \(f\). NoisyNet, specifically, is a stochastic
process parameterized by its weights and their standard deviation. We will define
this parameterized function as \(g_\phi\), and their sampling distribution as
\(q_\phi \in \mathcal{Q}\). Given a noise vector \(\xi\), we have
\(f(\bm{s}, \bm{a}) = g_\phi(\bm{a}, \bm{s}, \xi)\). This means that NoisyNet
gives a new policy function for each noise vector \(\xi\). Although this means
that NoisyNet can model stochastic policies, \citet{fortunato_noisy_2019} point
out that for the loss function it is optimising, there always exsit a deterministic
optimal policy. They show through emipirical analysis, however, that this does not
mean that the policy necessarily disregards the noise and converges to a deterministic
policy.

@@

\begin{equation}
    \KL[Q \Vert P] = \sup_{n \in \N, \X \in \mathcal{X}^n} \KL \left[ Q_\X \Vert P_\X \right],
\end{equation}

Now the stochastic processes are a neural network \(q_\phi\) and a prior \(p\). In a similar
manner to NoisyNet, we sample a function \(f\) from \(q\) by sampling a random noise vector
\(\xi\), and then defining \(f(\bm{x}) = g_\phi(\bm{x}, \xi)\). They further show that the
gradient of the KL-divergence for functions evaluated at the measurement set \(\X\) is

@@

\section{Method}
% Motivation:
% * Optimal regret
% * Meaningful priors
We will present a method based on functional variational Bayesian neural
networks~\citep{sun_functional_2019} that allows the following:

@@

@@

We would like to approximate the posterior distribution of the Q-value function.
\citet{osband_why_2017} prove that in finite horizon MDPs posterior sampling for
reinforcement learning has at least a near-optimal regret bound. They further conjecture
that this cound can be further improved to show an optimal regret bound. Additionally,
a posterior sampling based reinfrocement learning algorithm could be made to utilize
prior information or domain knowledge through an appropriate prior. This should improve
policy convergence rate.
Combined with the computational efficiency of posterior sampling, this motivates the
development of a Bayesian reinforcement learning algorithm with functional priors.

We can use the functional variational bayesian neural
network (FVBNN)~\citep{sun_functional_2019} framework discussed earlier.
NoisyNet uses one sample from Q for each optimization step. We will instead use
\(k\) samples from Q. This lets us add the KL-term from FVBNN to the loss function that
encourages stochasticity.
The maximization target becomes:
\begin{equation}
    \log p(\D_\omega^N \vert f) - \lambda \KL \left[ q(\f^{\D_\omega^N}, \f^\mathbf{M}) \Vert p(\f^{\D_\omega^N}, \f^\mathbf{M}) \right],
\end{equation}
where \(\f^{\D_\omega^N}\) is \(f\) applied to the dataset \(\D_\omega^N\), and \(\f^\mathbf{M}\) is \(f\) applied to
a set of i.i.d. random points \(\mathbf{M} = \{m_i \sim c \mid i=1,\dots,k\} \subseteq \state \times \action\)
with full support. We will let the measurement \(\mathbf{M}\) be the the set of all state-action
pairs for states we have already explored. If the MDP is ergodic, then \(\textbf{M}\) will eventually
have full support in \(\state \times \action\).

@@

\begin{table}
    \centering
    \caption{Hyperparameters for LFM.}\label{tab:hyperparams}
    \begin{tabular}{rl}
      \toprule % from booktabs package
      Hyperparameter & Value \\
      \midrule % from booktabs package
      \(k\) (number of samples) & 100\\
      \(J\) (Number of eigenvalues) & 6\\
      \(\lambda\) & \(\frac{1}{\lvert \mathcal{D}_s \rvert}\)\\
      \(\eta\) (Regularization coefficient for SSGE) & 0.95\\
      \bottomrule % from booktabs package
    \end{tabular}
\end{table}

\subsection{Uninformative Prior}
First, it is interesting to see how our method will compare to NoisyNet
in environments without the advantage of prior knowledge. The training curves
of our model compared to that of standard NoisyNet-DQN on the four selected OpenAI
Gym~\citep{brockman_openai_2016} enivornments is shown in Figure~\ref{fig:training_curves}.
We hypothesize that the added KL-term will push the algorithm keep weight uncertinaty high
in order to keep a higher uncertainty in the output. While both methods are able to
find an optimal policy in Cartpole, our method method uses considerably fewer frames
to do so.

@@

\cite{fortunato_noisy_2019} noted that the learned variance in their weight increased
in some games despite there existing an optimal deterministic solution, and the loss
giving no incentive to maintain any uncertainty. Figure~\ref{fig:mean_std} shows
the mean-absolute standard deviation for the penultimate and final layer. It is 
interesting to see that the standard deviation for NoisyNet continous to decrease
throughout the training, while LFM decreases faster, but then flattens out earlier.
This seems to indicate that it has found a policy that

@@

\section{Discussion}

\section{Conclusion}
We have developed a method for exploration with a DQN-style algorithm that can
effectively utilize domain knowledge for faster learning.

>>>>

%% Some suggested packages, as needed:
\usepackage{natbib} % has a nice set of citation styles and commands
    \bibliographystyle{plainnat}
    \renewcommand{\bibsection}{\subsubsection*{References}}
\usepackage{mathtools} % amsmath with fixes and additions
% \usepackage{siunitx} % for proper typesetting of numbers and units
\usepackage{booktabs} % commands to create good-looking tables
\usepackage{multicol}
\usepackage{tikz} % nice language for creating drawings and diagrams
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{bm}
\usepackage{subcaption}

@@

In this paper, we will introduce an extension to NoisyNet that puts it into a Bayesian context. A limitation of NoisyNet is that the initial uncertainty in the Q-value function is crucial to exploration. If the uncertainty is too high, algorithm will struggle to learn anything, while if the uncertainty is too low, there is nothing incentivizing exploration, and the algorithm will likely be stuck in a poor local minimum.

@@

\subsection{Noisy Networks}
In NoisyNet \citep{fortunato_noisy_2019}, the Q-function space can be seen
as a stochastic process on the probability space \((\Omega, \mathcal{F}, P)\).
The stochastic process can simply be written as
\(\{Q(\bm{s}, \bm{a}) : (\bm{s}, \bm{a}) \in \state \times \action \}\). For
any \(\omega \in \Omega\), \(Q(\cdot, \cdot, \omega)\) is a sample function
mapping \(\state \times \action \rightarrow \mathbb{R}\). To simplify notation
we will denote sample functions as \(f\). NoisyNet, specifically, is a stochastic
process parameterized by its weights and their standard deviation. We will define
this parameterized function as \(g_\phi\), and their sampling distribution as
\(q_\phi \in \mathcal{Q}\). Given a noise vector \(\xi\), we have
\(f(\bm{s}, \bm{a}) = g_\phi(\bm{a}, \bm{s}, \xi)\). This means that NoisyNet
gives a new policy function for each noise vector \(\xi\). Although this means
that NoisyNet can model stochastic policies, \citet{fortunato_noisy_2019} point
out that for the loss function it is optimising, there always exsists a deterministic
optimal policy. They show through emipirical analysis, however, that this does not
mean that the policy necessarily disregards the noise and converges to a deterministic
policy.

@@

\begin{equation}
    \KL[A \Vert B] = \sup_{n \in \N, \X \in \mathcal{X}^n} \KL \left[ A_\X \Vert B_\X \right],
\end{equation}

Now the stochastic processes is represented by a neural network \(q_\phi\) and a prior \(p\). In a similar
manner to NoisyNet, we sample a function \(f\) from \(q\) by sampling a random noise vector
\(\xi\), and then defining \(f(\bm{x}) = g_\phi(\bm{x}, \xi)\). They further show that the
gradient of the KL-divergence for functions evaluated at the measurement set \(\X\) is

@@

\section{Method}
% Motivation:
% * Optimal regret
% * Meaningful priors

\citet{osband_why_2017} prove that in finite horizon MDPs, posterior sampling for
reinforcement learning has at least a near-optimal regret bound. They further conjecture
that this can be improved to show an optimal regret bound. Additionally,
a posterior sampling-based reinforcement learning algorithm can be made to utilize
prior information or domain knowledge through an appropriate prior. This should improve
the policy convergence rate.
Combined with the computational efficiency of posterior sampling, this motivates the
development of a Bayesian reinforcement learning algorithm with functional priors.

We will present a method based on functional variational Bayesian neural
networks~\citep{sun_functional_2019} that allows the following:

@@

We would like to approximate the posterior distribution of the Q-value function.
%% BNN useful for both

@@

We can use the functional variational Bayesian neural
network (FVBNN)~\citep{sun_functional_2019} framework discussed earlier.
NoisyNet uses one sample from Q for each optimization step. We will instead use
\(k\) samples from Q. This lets us add the KL-term from FVBNN to the loss function that
encourages stochasticity.
The maximization target becomes:
\begin{equation}
    \log p(\D_\omega^N \vert f) - \lambda \KL \left[ q(\f^{\D_\omega^N}, \f^\mathbf{M}) \Vert p(\f^{\D_\omega^N}, \f^\mathbf{M}) \right],
\end{equation}
where \(\f^{\D_\omega^N}\) is \(f\) applied to the dataset \(\D_\omega^N\), and \(\f^\mathbf{M}\) is \(f\) applied to
a set of i.i.d. random points \(\mathbf{M} = \{m_i \sim c \mid i=1,\dots,k\} \subseteq \state \times \action\)
with full support. We will let the measurement \(\mathbf{M}\) be the the set of all state-action
pairs for states we have already explored. If the MDP is ergodic, then \(\textbf{M}\) will eventually
have full support in \(\state \times \action\).

@@

\begin{table}
    \centering
    \caption{Hyperparameters for LFM.}\label{tab:hyperparams}
    \begin{tabular}{rll}
      \toprule % from booktabs package
      \multicolumn{2}{c}{Hyperparameter} & Value \\
      \midrule % from booktabs package
      \(k\)         & Number of function samples & 100\\
      \(J\)         & Number of eigenvectors & 6\\
      \(\lambda\)   & KL regularization parameter & \(\frac{1}{\lvert \mathcal{D}_s \rvert}\)\\
      \(\eta\)      & Regularization coefficient for SSGE & 0.95\\
      \bottomrule % from booktabs package
    \end{tabular}
\end{table}

\subsection{Uninformative Prior}
First, it is interesting to see how our method will compare to NoisyNet
in environments without the advantage of prior knowledge. The training curves
of our model compared to that of standard NoisyNet-DQN on the four selected OpenAI
Gym~\citep{brockman_openai_2016} enivornments is shown in Figure~\ref{fig:training_curves}.
We hypothesize that the added KL-term will push the algorithm keep weight uncertinaty high
in order to keep a higher uncertainty in the output. While both methods are able to
find an optimal policy in Cartpole, our method uses considerably fewer frames
to do so.

Future

@@

\cite{fortunato_noisy_2019} noted that the learned variance in their weight increased
in some games despite there existing an optimal deterministic solution, and the loss
gives no incentive to maintain any uncertainty. Figure~\ref{fig:mean_std} shows
the mean-absolute standard deviation for the penultimate and final layer. It is 
interesting to see that the standard deviation for NoisyNet continous to decrease
throughout the training, while LFM decreases faster, but then flattens out earlier.
This seems to indicate that it has found a stable policy, where optimizing further would
be overfitting to noise.

@@

\section{Conclusion and Discussion}
We have presented a method for exploration with a DQN-style algorithm that can
effectively utilize domain knowledge for faster learning. Our results show that
it outperforms standard DQN with NoisyNet exploration with regards to sample efficiency
on all four selected environments even without prior knowledge.

One interesting avenue for future work is to extend the approach to methods other than
the standard DQN. A functional Bayesian approach for policy evaluation would permit direct
prior distribution in policy space rather than in value space, which seems like a
more intuitive prior in many domains.
